Fréchet Distance

### 2022-12-01 docs/paper.tex
Algorithm for discrete Fréchet distance

\section{Algorithms}
\label{sec_algorithms}
% Chapter summary and structure
Several algorithms to compute the Fréchet distance of polygonal curves were proposed, 
starting from the Original algorithm for the continuous variant by Alt and Godau in 1995 \cite{alt_computing_1995}.
The algorithm is based on a concept called free-space diagram and gives a baseline for newer improved versions, 
e.g., the version by Buchin et al. \cite{buchin_four_2017}. 
This chapter is structured as follows: 
first the original algorithm for the continuous Fréchet distance is described in \autoref{sec_original_algorithm},
then in \autoref{sec_improved_algorithm} a more recent improved algorithm for the same problem is given.
Afterwards, the modification necessary to compute the weak variant are described in \autoref{sec_improved_algorithm}.
Finally, an algorithm to compute the discrete Fréchet distance is given in \autoref{sec_discrete_algorithm}.
% Finally

@@

\subsection{Improved Version?}
\label{sec_improved_algorithm}

% Algorithms for different variants
\subsection{Algorithm for weak Fréchet distance}
\label{sec_weak_algorithm}
The weak Fréchet distance, as described in \autoref{sec_weak_frechet_distance},
the computation becomes less complex compared to the continuous Fréchet distance.
Specifically, we can ignore critical values of the third type, 
since it is no longer required to move monotonically 
and one could just go left or down in the free space diagram taking any arbitrary continuous path.
Therefore, the focus shifts towards critical values of the second type, 
those where new passages between neighboring cells open up.
Alt and Godau \cite{alt_computing_1995} proposed two algorithms to compute the weak Fréchet distance.

@@

\subsection{Algorithm for discrete Fréchet distance}
\label{sec_discrete_algorithm}

>>>>

\section{Algorithms}
\label{sec_algorithms}
% Chapter summary and structure
Several algorithms to compute the Fréchet distance of polygonal curves were proposed, 
starting from the Original algorithm for the continuous variant by Alt and Godau in 1995 \cite{alt_computing_1995}.
The algorithm is based on a concept called free-space diagram and gives a baseline for newer improved versions, 
e.g., the version by Buchin et al. \cite{buchin_four_2017}. 
This chapter is structured as follows: 
first the original algorithm for the continuous Fréchet distance is described in \autoref{sec_original_algorithm},
then in \autoref{sec_improved_algorithm} a more recent improved algorithm for the same problem is given.
Afterwards, the modification necessary to compute the weak variant are described in \autoref{sec_weak_algorithm}.
Finally, an algorithm to compute the discrete Fréchet distance is given in \autoref{sec_discrete_algorithm}.
% Finally

@@

\subsection{Improved Version?}
\label{sec_improved_algorithm}
\cite{van_leusden_novel_2013}
\cite{buchin_four_2017}

% Algorithms for different variants
\subsection{Algorithm for the Weak Fréchet Distance}
\label{sec_weak_algorithm}
The weak Fréchet distance, as described in \autoref{sec_weak_frechet_distance},
the computation becomes less complex compared to the continuous Fréchet distance.
Specifically, we can ignore critical values of the third type, 
since it is no longer required to move monotonically 
and one could just go left or down in the free space diagram taking any arbitrary continuous path.
Therefore, the focus shifts towards critical values of the second type, 
those where new passages between neighboring cells open up.
Alt and Godau \cite{alt_computing_1995} proposed two algorithms to compute the weak Fréchet distance.

@@

\subsection{Algorithm for the Discrete Fréchet Distance}
\label{sec_discrete_algorithm}
The computation of the discrete Fréchet distance, as defined in \autoref{def_discrete_frechet_distance}, is very easy and straight forward 
using a dynamic programming approach as originally suggested by Eiter et al. \cite{eiter_computing_1994}.
Recall that the polygonal curves $P = (u_1, \dots, u_p)$ and $Q = (v_1, \dots, v_p)$ are defined as sequences of points.
The algorithm works by recursively computing the minimal required $\varepsilon$ for partial curves 
$P' = (u_1, \dots, u_i)$ and $Q' = (v_1, \dots, v_j)$
using $c(i,j)$ as defined in \autoref{def_recursive_forumla_discrete_frechet_distance}.
Starting from $c(1,1)$, which is the distance between the starting points $u_1$ and $v_1$,
all other values of $c(i,j)$ are computed iteratively in a bottom up manner 
and memoization on the previously computed values of $c(i,j)$.
The resulting algorithm has runtime of $\mathcal{O}(pq)$
because the constant time function $c(i,j)$ is called $pq$ times \cite{eiter_computing_1994}.

\begin{mydef}
  \label{def_recursive_forumla_discrete_frechet_distance}
  The discrete Fréchet distance of polygonal curves $P' = (u_1, \dots, u_i)$ and $Q' = (v_1, \dots, v_j)$
  can be computed by
  \begin{align*}
    c(i,j) &= \begin{cases}
      dist(u_1, v_1), &\text{if } i=1 \text{and } j=1 \\
      \max \{c(i-1,j), dist(u_i, v_1)\}, &\text{if } i>1 \text{and } j=1 \\
      \max \{c(i,j-1), dist(u_1, v_j)\}, &\text{if } i=1 \text{and } j>1 \\
      \max \{c(i-1,j), c(i,j-1), c(i-1,j-1), dist(u_i, v_j)\}, &\text{if } i>1 \text{and } j>1
    \end{cases} \\
    &= \delta_{dF}(P',Q')
  \end{align*}
\end{mydef}
